---
title: "PackageIntroduction"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{PackageIntroduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

library(lemon)
knit_print.data.frame <- lemon_print
```


```{r setup}
library(dataValidation)
library(tidyselect)
```

# Distinct

## Confirm Distinct

In data analysis tasks we often have data sets with multiple possible ID columns, but it's not always clear which combination uniquely identifies each row.

sample_data1 has 125 row with 3 ID type columns and 3 value columns.

```{r render=lemon_print}
head(sample_data1)
```
Let's use `confirm_distinct` iteratively to find the uniquely identifying columns of sample_data1.

```{r}
sample_data1 %>% 
  confirm_distinct(ID_COL1)
```

```{r}
sample_data1 %>% 
  confirm_distinct(ID_COL1, ID_COL2)
```

```{r}
sample_data1 %>% 
  confirm_distinct(ID_COL1, ID_COL2, ID_COL3)
```

Here we can conclude that the combination of 3 ID columns is the primary key for the data.

## Determine Distinct

These steps can be automated with the wrapper function `determine distinct`. 

```{r}
sample_data1 %>% 
  determine_distinct(matches("ID"))
```

# Cardinality

In data-modeling cardinality refers to the mapping between the rows of two tables, see [Wikipedia: cardinality](https://en.wikipedia.org/wiki/Cardinality_(data_modeling))  

`confirm_cardinality` tells you the mapping between two columns in a data frame:

- 1 - 1 mapping
- 1 - many mapping
- many - 1 mapping
- many - many mapping


## Confirm cardinality

`confirm_cardinality` gives the option to view which type of mapping is associated with each individual row. 

```{r}
sample_data1 %>% 
  confirm_cardinality(ID_COL1, ID_COL2, view = F)
```

## Determine cardinality

```{r}
sample_data1 %>% 
  determine_cardinality(everything())
```

# Overlap

The `overlap` functions give a venn style description of the values in 2 columns. This is especially useful before performing a `join` function, and you want to confirm that the dataframes have matching keys. 

## Confirm Overlap

`confirm_overlap` is different from the other `confirm` functions in that it takes 2 vectors as arguments, instead of a data frame. This is to allow the user to test overlap between different dataframes, or arbitrary vectors if necesary. 

```{r render=lemon_print}

confirm_overlap(iris$Sepal.Width, iris$Petal.Length)
```

## Determine Overlap

`determine_overlap` takes a dataframe and a tidyselect specification, and returns a tibble summarizing all of the pairwise overlaps. Only pairs with matching types are tested. 

```{r render=lemon_print}
iris %>% 
  determine_overlap(everything())
```
Note that the `overlap` functions only test pairwise overlaps. For multi-column and large-scale overlap testing, see [Complex Upset Plots](https://krassowski.github.io/complex-upset/)
