---
title: "PackageIntroduction"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{PackageIntroduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(dataValidation)
library(tidyselect)
```

## Confirm Distinct

In data analysis tasks we often have data sets with multiple possible ID columns, but it's not always clear which combination uniquely identifies each row.

sample_data1 has 3 ID type columns and 3 value columns.

```{r}
DT::datatable(sample_data1)
```
Let's use `confirm_distinct` iteratively to find the uniquely identifying columns of sample_data1.

```{r}
sample_data1 %>% 
  confirm_distinct(ID_COL1)
```

```{r}
sample_data1 %>% 
  confirm_distinct(ID_COL1, ID_COL2)
```

```{r}
sample_data1 %>% 
  confirm_distinct(ID_COL1, ID_COL2, ID_COL3)
```

Here we can conclude that the combination of 3 ID columns is the primary key for the data.

## Determine Distinct

These steps can be automated with the wrapper function `determine distinct`. 

```{r}
sample_data1 %>% 
  determine_distinct(matches("ID"))
```

## Confirm Mapping

```{r}
sample_data1 %>% 
  confirm_mapping(ID_COL1, ID_COL2, view = F)
```

## Determine Mapping

```{r}
sample_data1 %>% 
  determine_mapping(everything())
```


